Hyper-linked Communications: WebRTC enabled asynchronous collaboration

### 2017-02-22 artigo/paper.tex
A few english revisions.

From an user experience point of view, we designed our collaboration application around the concept of group conferences (or conference rooms).
Users can create group conferences and invite other users to participate in these, either by adding the other user or by sharing a \gls{URL} (invitation link).
In order to meet user expectation, a group conference provides functionality similar to those found in established applications such as \emph{Skype} or \emph{Google Hangouts}.
Among other common features, one can view a video composed by all participants or a video of a selected participant.
Each participant may use more than one camera, in order to show different angles or objects, e.g. a white board and himself.
The source of the video can be selected manually or automatically by the talking participant.
Participating users may choose to view only, share their microphone, camera or screen.

Besides the videos, a group conference includes a chat, collaborative text editor, list of users and document sharing.
The major differentiating feature is a time-line bar, for navigating in time.
While the group conference is taking place, this time bar is continuously progressing as time goes by.
The user is free to slide over to any instant in the past or back to real-time.
Changing the instant of time updates the state of not only the video but of other collaboration elements. % ("other collaboration elements", nao foi implementado, apenas para conteudo sobreposto...) %

@@

In order to limit the processing requirement in the client application (browser), we decided to use a server to combine the audio and video streams from all the participants into a single video and a single audio track. This techinique is called \gls{MCU}.
This also helps conserve battery on mobile devices.
Each user listens to the combined audio track and may choose to see a single participant or all of them, in this case using the combined video track.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/modules2.png}
	\caption{System Modules}
    \label{fig:modules}
\end{figure}

@@

In order to simplify our solution, the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules are deployed within the same server, so to facilitate deployment.
To this set of modules, we call \emph{backend}.
The \emph{Application} module is split between client and server part.
The \emph{Signaling} module is implemented together with the application server.

@@

By default, \emph{Kurento Repository}, used for storing the recorded video and audio streams, is implemented over \emph{MongoDB}.
For convenience, our storage model also uses the same database.
In order to offer all the functionality, some information about objects must be stored on database, such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.

@@

\resizebox{\columnwidth}{!}{%
    \begin{tabular}{cccccccc@{}m{0pt}@{}}
	\hline 
	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[2pt]\cline{1-5}\cline{7-7}
	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[2pt]\hline
	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[2pt]\hline
	\end{tabular}
}
\end{table}

@@

The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room. 
It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.

@@

	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{figures/devices2.png}
		\caption{Multiple devices per user}
		\label{fig:devices}
	\end{figure}

@@

We tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
We have tested as well the performance of our solution by measuring the used resources.
Those performance tests are crucial to ensure that our solution is in fact stable and scalable.

@@

Figure \ref{fig:test_client_net} shows the same events from the respective network usage on the client side.
We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS} and maintains that upload rate throughout.
Whenever a new client enters the conference, we observe that \gls{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room.
After a while, \gls{KMS} realizes that the network can handle the increase of clients and sends the video with a better quality to every participant.
When a user leaves the conference room \gls{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
As expected, the maximum bandwidth used is constant and independent from the amount of participants.

>>>>

From an user experience point of view, we designed our collaboration application around the concept of group conferences (or conference rooms).
Users can create group conferences and invite other users to participate in these, either by adding the other user or by sharing a \gls{URL} (invitation link).
In order to meet user expectation, a group conference provides functionality similar to those found in established applications such as \emph{Skype} or \emph{Google Hangouts}.
Among other common features, one can view a video composed by all participants or a video of a selected participant.
Each participant may use more than one camera, in order to show different angles or objects, e.g. a white board and himself.
The source of the video can be selected manually or be automatically assigned to the talking participant.
Participating users may choose to view only, share their microphone, camera or screen.

Besides the videos, a group conference includes a chat, collaborative text editor, list of users and document sharing.
The major differentiating feature is a time-line bar, for navigating in time.
While the group conference is taking place, this time bar is continuously progressing as time goes by.
The user is free to slide over to any instant in the past or back to real-time.
Changing the instant of time updates the state of not only the video but also of other collaboration elements. % ("other collaboration elements", nao foi implementado, apenas para conteudo sobreposto...) %

@@

In order to limit the processing requirement in the client application (browser), we decided to use a server to combine the audio and video streams from all the participants into a single video and a single audio track. This technique is called \gls{MCU}.
This also helps conserve battery on mobile devices.
Each user listens to the combined audio track and may choose to see a single participant or all of them, in this case using the combined video track.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/modules.pdf}
	\caption{System Modules}
    \label{fig:modules}
\end{figure}

@@

In order to simplify our solution, the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules are deployed within the same server, as to facilitate deployment.
To this set of modules, we call \emph{backend}.
The \emph{Application} module is split between client and server part.
The \emph{Signaling} module is implemented together with the application server.

@@

By default, \emph{Kurento Repository}, used for storing the recorded video and audio streams, is implemented over \emph{MongoDB}.
For convenience, our storage model also uses the same database.
In order to offer all the functionality, some information about objects must be stored on the database, such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.

@@

\resizebox{\columnwidth}{!}{%
    \begin{tabular}{cccccccc@{}m{0pt}@{}}
	\hline 
	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
	\end{tabular}
}
\end{table}

@@

The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room. 
It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.

@@

	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{figures/devices.png}
		\caption{Multiple devices per user}
		\label{fig:devices}
	\end{figure}

@@

We tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
We have also tested the performance of our solution by measuring the used resources.
Those performance tests are crucial to ensure that our solution is in fact stable and scalable.

@@

Figure \ref{fig:test_client_net} shows the same events regarding network usage from the perspective of the client.
We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS} and maintains that upload rate throughout.
Whenever a new client enters the conference, we observe that \gls{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room.
After a while, \gls{KMS} realizes that the network can handle the increase of clients and sends the video with a better quality to every participant.
When a user leaves the conference room \gls{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
As expected, the maximum bandwidth used is constant and independent from the amount of participants.
